\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-28 metatheory-condensed.tex
W-types

We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
is inhabited by the \emph{codes} of types, with $\El{t}$ decoding~$t$ and
$\codety{T}$ encoding type~$T$ (rules \ruleref{ty/el} and \ruleref{tm/c}).
MLTT has an infinite hierarchy of universes;
we omit universe levels for conciseness.

@@

    (* s[@₁@]  : {} | {} *)
    - o[@₁@]  = μ+ {} s[@₁@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]}
    (* s[@₂@]  : {ty : 𝕎τ[@₁@]} | [ty : 𝕌; bool : ty; ...] *) 
    - o[@₂@]  = μ+ o[@₁@] s[@₂@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
    (* s[@₃@][@₁@]  : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]} 
            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] (μ+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
    (* s[@₃@] : {ty : 𝕎(τ[@₁@]); tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
          | [ty : 𝕌; ... ; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)] *)
    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}

@@

Each hiding comment corresponds to the textual-aligned sealing derivation on RHS. For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested sigma type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.

@@

Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) sigma type instead of a linkage type. For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$. Because the latter cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type. For simplicity, we make sealing a term in the context of compilation type instead of linkage type.

>>>>

We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
is inhabited by the \emph{codes} of types,
with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
and $\codety{T}$ encoding type~$T$ (rule \ruleref{tm/c}).
MLTT has an infinite hierarchy of universes;
we omit universe levels in the presentation for conciseness.

\noindentparagraph{W-types.}

W-types~\cite{martin1984intuitionistic} are a standard way to model
inductive types in MLTT.
A constructor of some inductive type~$T$ is given by a pair of types
$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
it constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
As the rules suggest, our formulation additionally supports
\emph{W-type signatures} (judgment form \goodWSig{\Gm}{\wsig}{n})
composed of $n$ pairs of types, each of which represents a constructor.

@@

    (* s[@₁@]  : {} | {} *)
    - o[@₁@]  = μ+ {} s[@₁@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]}
    (* s[@₂@]  : {ty : 𝕎τ[@₁@]} | [ty : 𝕌; bool : ty; ...] *) 
    - o[@₂@]  = μ+ o[@₁@] s[@₂@] ⋆ : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
    (* s[@₃@][@₁@]  : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]}
            | [ty : 𝕌; ... ; tm : 𝕌; tvar : tm; ...] *)
    - o[@₃@][@₁@]  = μ+ o[@₂@] s[@₃@] ([@μ@]+ ..) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)}
    (* s[@₃@] : {ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)} 
          | [ty : 𝕌; ... ; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig)] *)
    - o[@₃@] = μ+ o[@₃@][@₁@] s[@₃@] (λ (Wrec ..)) : 𝕃{ty : 𝕎τ[@₁@]; tm : 𝕎τ[@₂@]; subst' : 𝕃(RecSig); subst : tm → nat → tm → tm}

@@

Each hiding comment corresponds to the textual-aligned sealing derivation on RHS. For legibility, curly brackets indicate a signature (omitting sealing at type level for readability) and square bracket indicate a nested $\Sigma$-type, corresponding to sealing judgment $\goodSeal{}{s}{\sigma}{A}$.

@@

Due to the definition of sealing, we can see the importance of compilation type, as the result of a sealing is usually a (nested) $\Sigma$-type instead of a linkage type. For example, $\TyLkg{\{a : S(\bot), b : a \to \bot\}}$ can only be sealed into $\TyTkg{\{a : \cU, b : a \to \bot\}}$ but never $\TyLkg{\{a : \cU, b : a \to \bot\}}$. Because the latter cannot be inhabitted---by projecting the second field of the $ t : \TyLkg{\{a : \cU, b : a \to \bot\}}$, we can get a proof of bottom type. For simplicity, we make sealing a term in the context of compilation type instead of linkage type.
